---
title: "Homework 6 Function in R"
author: "Frederik Normann Holm"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## installing and loading needed packages for the exercises

```{r packages}
library("gapminder")
library(dplyr)
```

## testing Gapminder command

Testing to see if it loaded properly
```{r Gapminder, echo=FALSE}
gapminder
```

## Exercise 1. 

Writing a defensive function that calculates the GDP of a given country in the gapminder dataset
Choosing to go with Norway for this example

```{r def function}
calculateGDP <- function(dat, year=NULL, country=NULL) {
  if(!is.null(year)) {
    dat <- dat[dat$year %in% year, ]
  }
  if (!is.null(country)) {
  dat <- dat[dat$country %in% country,]
  }
  gdp <- dat$pop * dat$gdpPercap
  
  new <- cbind(dat, gdp=gdp)
  return(new)
}

calculateGDP(gapminder, country="Norway")
```

Using the same defensive function to calculate the GDP of Denmark in 1967, 1977, 1987, 1997, 2007, and 2017.

```{r denmark }
calculateGDP(gapminder, year=c(1967,1977,1987,1997,2007,2017), country="Denmark")
```

## Exercise 2

Creating the script that loops over every country in the dataset as well as checking for B's and life expectancy
in order to get every country that starts with a B in the gapminder dataset i make use of the grep function

```{r life exp}
grep("^B", unique(gapminder$country), value = TRUE)

lowerThreshold <- 50
upperThreshold <- 70
Bcountries <- grep("^B", unique(gapminder$country), value = TRUE)

for(iCountry in unique(gapminder$country)){
  tmp <- mean(subset(gapminder, country==iCountry)$lifeExp)
  
  if(tmp < lowerThreshold){
    cat("Average Life Expectancy in", iCountry, "is less than", lowerThreshold, "\n")
  }
  else if(tmp > lowerThreshold && tmp < upperThreshold){
    cat("Average Life Expectancy in", iCountry, "is between", lowerThreshold, "and", upperThreshold, "\n")
  }
  else{
    cat("Average Life Expectancy in", iCountry, "is greater than", upperThreshold, "\n")
  }
  rm(tmp)
}
```
For this exercise, i had trouble including the grep(^"B") function in my loop so i could never exclusively 
print the countries that started with B and had life expectancies under 50, between 50 and 70, and over 70
