---
title: "CDS Homework 7 - practice web scraping"
author: "Frederik Normann Holm"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading necessary packages
I have chosen to scrape records and statistics of my favorite football team AS Roma. 
Firstly, i load a bunch of helpful packages (including rvest) from the library
```{r libraries, warning=FALSE, message=FALSE}
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(janitor)
```

## Scraping the data from Wikipedia
now i scrape the website by submitting the url 
```{r url}
url <- "https://en.wikipedia.org/wiki/List_of_A.S._Roma_records_and_statistics"
url_html <- read_html(url)
```

```{r scrape-table}
whole_table <- url_html %>% 
 html_nodes("table") %>%
 html_table()  
str(whole_table)
```
Unlisting and combining as colums
```{r html-to-df}
new_table <- do.call(cbind,unlist(whole_table, recursive = FALSE))
head(new_table)
```
## Scraping only the wanted tables from Wikipedia page
By telling r that i only want the first 3 tables on the Wikipedia page, i put a square bracket after the "whole_table".
```{r scrape-funtion}
scrape_roma_records <- function(website){
	url <- read_html(website)
	whole_table[1:3] <- url %>% 
		html_nodes("table") %>%
		html_table()  # result is a list
	whole_table[1:3] <- do.call(cbind,unlist(whole_table[1:3], recursive = FALSE))
}
```
Testing that the function works on the clubs records that i selected it to scrape
```{r testing-function}
tablegoals <- scrape_roma_records("https://en.wikipedia.org/wiki/List_of_A.S._Roma_records_and_statistics")
tablegoals %>% 
	print(tablegoals)
```


